Unsupervised Semantic Mapping for Healthcare Data Storage Schema

### 2021-06-04 ManuScript/IEEEAccess/ManuScript.tex
final updates

\begin{keywords}
\sout{Electronic medical records, Health information management, Medical information systems,  Text processing, Unsupervised learning}
\end{keywords}

@@

@@

Data interoperability resolves the heterogeneity between data artificats to enable seamless and interpretable communcation among source and target organizations, while preserving the data's original intention during storage, communication, and usage (as defined by IEEE 610.12 \cite{geraci1991ieee}, Health Level Seven International (HL7), and Healthcare Information and Management Systems Society HIMSS \cite{Himss_url2013}).
On the other hand, process interoperability regulates the communication among organizational processes to provide compatability between process artifacts within and seamless transformations across different organizations\cite{khan2013process}. Lastly, knowledge interoperability provides a sharing mechanism for reusing interpretable medical knowledge, acquired through expert intervention and other mechanisms, across decision support systems \cite{ali2017multi}.

@@

Healthcare Standards such as HL7 - Fast Healthcare Interoperability Resources (FHIR), and openEHR provide the foundations for storing and communicating medical data, through the use of well defined protocols. While Systematized Nomenclature of Medicine—Clinical Terms (Snomed-CT) \cite{snomedct_url} and Logical Observation Identifiers Names and Codes (LOINC) \cite{loinc} provide a standard definition for clinical terminologies and laboratory tests, respectively. Similarly Medical Logic Module (MLM) provides a standardized way for expressing medical knowledge. However, the plethora of standards, necessitates the creation of bridging standards, that can resolve the heterogeniety between the medical standards. Substantial effort has gone into this endeavor with the Clinical Information Modeling Initiative (CIMI) \cite{CIMI2015} taking the lead in bridging the gap between HL7v3 and openEHR. Similarly, SNOMED CT and LOINC are working to resolve the redundancies between the two terminological standards since 2013. This healthcare interoperability solution follows a formal, albiet long process, which is greatly dependent on the human factor. However, the current healthcare scenario, requires a quick solution to create a scaffolding of an interoperable bridge between various healthcare providers. It is also important to ensure that this scaffolding should be able to support the formal standardization processes of the future. In \cite{Satti2020}we have presented the Ubiquitous Health Platform (UHP), which provides semantic reconciliation-on-read based data curation for resolving data interoperability between various schema. This methodology is based on the creation and management of schema maps, that can provide the framework for transforming a source schema into a target schema. 
In the current manuscript, we will present our research work to build and manage the schema map knowledge base. Overall, our methodology is based on the creation, evaluation, and application of a novel schema matching technique to identify the relationships between attributes of the participating medical data schema. Since the terms used to identify attributes in the data schemas are not defined in any standard way, it is important to first identify the component words of the attribute term and then to append semantic concepts with these to create a meaningful sentence. This process is based on word expansion using concept lookup from Unified Medical Language System (UMLS). Once the sentence has been created, it is then trivial to create its embedded vectors representation using transformer based pre-trained models. The cosine similarity of any two embedded vectors can then indicate the degree of similarity between the original attributes. The motivation behind this methodology along with its development and evaluation details will be presented in the following flow:

@@

@@

\section{Methodology}
\label{methodology}
Healthcare interoperability, with a focus on non-standard compliant medical schema, is dependent on the generation and validation of schema maps, as discussed above. To this end, the creation of a cohesive workflow is of utmost importance. In our earlier work \cite{Satti2020} we used maximum sequence identification and suffix tress based matching for syntactic matching of two distinct data schemas. This was followed by semantic concept enrichment and subsequently concept matching, for creating rules in the form of schema maps. The simplified mapping functions, thereby created, provided a simple methodology for converting semi-structured medical data, into an interpretable, model form.
In our current methodology we have utilized state-of-the-art natural language processing (NLP) techniques to extract the schema mapping rules from semi-structured data schemas. This methodology is based on identifying similarity between vector representations of two attributes, belonging to different medical schemas. 
Traditional NLP techniques such as WordNet are able to convert a word into an embedded vector, while Bidirectional Encoder Representations from Transformers (BERT) extracts an embedded vector from a sentence \cite{zhu2018exploring}. However, the terms forming the attribute names are bigger than a word (combination of multiple words) and smaller than a sentence. In order to resolve this problem, we extracted the set of suffixes from the terms forming the attribute names. The bidirectional nature of BERT, allows the creation of contextual embedded vectors, where each target word is affected by its neighboring words. Hence to convert the set of suffixes into a sentence, we collected the set of concepts corresponding to each suffix, from UMLS. This operation has two effects, firstly it is used to remove any suffix, which does not have a corresponding concept and secondly the extracted concepts are used to add context to each suffix and produce a contextual sentence. 
The following subsections provide the practical details for our methodology, from schema acquisition, to attribute name expansion, and finally schema map generation.

\subsection{Schema acquisition}
\label{schema_acq}
In the first step of our semantic reconciliation methodology, we simulate medical data acquisition from five distinct Electronic Medical Records (EMR) storage systems ($S$). These include patient reports from OpenEMR ($s_1$), 100,000 patient records from EMRBOTS ($s_2$) \cite{kartoun2016methodology}, custom database design by Pan et. al($s_3$) for supporting regional clinics and health care centers in China \cite{pan2016design}, clinical knowledge discovery tool MedTAKMI-CDI ($s_4$) \cite{inokuchi2007medtakmi}, and our custom implementation ($s_5$). Each of these medical systems as shown in \ref{fig:schema}, follows the relational database design, with logical entities, such as demographics, diagnosis, medicine or others, placed into tables which can be further linked to one or more tables. While the database design implemented by each of these systems, fulfills the need of their respective information processing applications, the lack of interoperability, in terms of identifying similar attributes or exchanging the medical data is very much evident here.

A similar notion of data heterogeneity, in terms of medical data schema, is evident across the healthcare domain. This is caused by various factors, including the lack of one all-encompassing, and universally applicable terminological standard and different normalization level for representing attributes.\\
In the former case, while SNOMED-CT provides a mechanism for identifying the standard codes for clinical terms and LOINC can be used for laboratory related terms, most attribute names are created based on the gut feeling of the database designer. Additionally, while these codes can be used to represent data instances, the data schema, achieves no benefit from the same.  Consider the terms ``name'' and ``patientName'', which refer to the same attribute of the patient entity.  However, since there is no standard way to represent this attribute, both are considered correct ($s_1$ and $s_3$ use the former representation, while $s_2$ and $s_5$ use the latter).\\
In the later case, differences in normalization cause semantic differences, due to which some data could be available in one schema but absent in others, such as OpenEMR demographics identifying the patient's residential location using specific attributes like ``Address'', ``City'', ``State'', ``Postal Code'', ``Country'', and others. Similarly, ``EncounterDate'' from $s_5$ is semantically similar to ``BeginDate'' of ``openemr\_MedicalProblems'' table in $s_1$, ``AdmissionStartDate'' of ``LabsCorePopulatedTable'' in $s_2$, 
``time'' in ``Diagnosis'' table of $s_3$, and  ``dateOfAdmission'' in ``Diagnosis'' and ``CareHistory'' tables of $s_4$.
Finally, $s_1$ and $s_3$ have separate tables containing the medicinal prescription details, however the same details are unavailable in $s_2$, $s_4$, $s_5$. Once again, this is not an incorrect behavior since this information, might not be a part of the context or the requirements for the EMR/EHR storage systems.
In fact, the change in context of the medical data storage system from the initial time of development to a later stage of collaborative processing systems, is the main cause of heterogeneity. In order to provide an interoperable solution, it is therefore necessary to enhance the semantics of each data attribute by its contextually equivalent sentence.

\begin{figure*}[tb!]
	\centering
	\includegraphics[scale=0.53]{suffixAttribute}
	\caption{Suffix attributes.}
	\label{fig:suffix_attribute}
\end{figure*}

\subsection{From Attribute to Sentence}
\label{schema_prepro}
In order to process the EMR/EHR schema set $S$ and produce a set of corresponding semantically enriched sentences, we use the data representation $s_i$, generated through the process explained in sequence acquisition to collect the various medical fragments in memory. We then iterate over these fragments, building a set of attributes, distinguished by their name, schema's name, table's name, schema's version, source, and recorded data. This entails that ``PatientID'' from each of the four tables in $s_2$, and ``patientID'' from five tables in $s_4$, would result into nine attributes (assuming, as in the current case, of no differences in versions of these systems).
For each attribute, we then generate the suffix array, which provides all possible substring representations contained within the attribute name. In order to generate the set of suffixes, we employ three strategies, forward suffix generation, whereby for a word $w$ of length $n$, $n-1$ suffixes of size 2 to $n-1$ are produced, backward suffix generation, to produce $n-1$ suffixes in reverse order with size $n-1$ to 2, and regular expression based suffix generation, which splits each word on, change of case, special characters(such as -, \_, !, and others), and numbers. In this way a large list of suffixes is generated, which is combined using a ``TreeSet'' data structure of Java, which internally sorts this list as well. An example of this suffix generation process, using the attribute name ``dateOfAdmission'' as it appears in $s_4$ is shown in \ref{fig:suffix_attribute}.

Suffix strings for similar attributes such as ``AdmissionStartDate''. ``diseaseNameOnAdmission'', and ``AdmissionEndDate'' appear in $s_2$, produce many, synctactically similar suffixes, to the presented example. This process, is only able to generate syntactic suffixes, producing many incoherent and unrelated suffixes. In order to counter this problem, and to limit the list of suffixes within the domain, we then query UMLS, with exact search strategy, looking for the existence of any concepts, against each suffix. In case, no semantic concept is found for a particular suffix, it is removed from the final Suffix Array. On the other hand, if atleast one semantic concept is found against the queried suffix, it is retained. Meanwhile the process continues for the next attribute, then the next table, and finally the next system, till no further processing is possible. 
The set of suffixes and their corresponding concepts are then used to build the sentence, where by each concept, corresponding to a suffix is appended next to the suffix. An example of the resultant sentence for the attribute ``DateOfAdmission'' is shown in Table \ref{tab:sentence_example}.

\begin{table}[]
	\centering
	\caption{Sentence created from the attribute name ``DateOfAdmission''}
	\label{tab:sentence_example}
	\begin{tabular}{p{\linewidth}}
		Date Value type - Date date allergenic extract Date in time Data types - Date Date Fruit;Of SPI1 wt Allele SPI1 gene TAF1 wt Allele BRIP1 gene Within Degrees fahrenheit Oral contraception BRIP1 wt Allele;Da Displacement of abomasum dalton Anterior descending branch of left coronary artery deca units cytarabine/daunorubicin protocol Dai Chinese Asymptomatic diagnosis of Drug Accountability Domain;ion Iontophoresis Route of Drug Administration Ions;on SPARC wt Allele Osteonectin SPARC gene On (qualifier value) Upon - dosing instruction fragment;Admission Admission activity Hospital admission;Dat SLC6A3 gene SLC6A3 wt Allele dopamine transporter Direct Coombs test SLC6A3 protein, human Test Date cytarabine/daunorubicin/thioguanine Alzheimer's Disease;mission Religious Missions;
	\end{tabular}
\end{table}

@@

Schema Maps, provide an interoperable bridge between two medical systems ($s_i \wedge s_j$), by identifying the semantic relationship between their participating attributes. This identification is based on the similarity between the embedded vectors, of the semantically enriched sentences, corresponding to each data attribute. While the embedded vectors can be generated using any methodology (we tested 11 methodologies, with WordNet and 10 models based on BERT (including BERT base, BERT large, distilBert, and RoBERTa) further detailed in section \ref{experimentalSetup}), the large/STSb version of RoBERTa\cite{reimers-2019-sentence-bert},provides the best results. The pair of embedded vectors thus produced are then used to calculate cosine similarity, which is based on the inverse cosine distance between them. For our classification, we used the raw results (unnormalized) of cosine similarity, which produces a score between -1, and 1. Cosine similarity score of 0 indicates orthogonal relationship between the two vectors, which in our scenario indicates that the two sentences, and by extension their attributes are not related to each other. -1 indicates, inverse relationship between the attributes, while 1 indicates the two attributes are very much the same.
For producing our schema maps, we are interested in three types of relationships, ``equal'' (the two attributes are same), ``related'' (the two attributes are related to each other), and ``unrelated'' (no relationship between the attributes). In order to classify the similarity results, into one of these three classes, we then calculated the best thresholds, using Matthews Correlation coefficient (MCC) \cite{chicco2020advantages} for classifying each instance as ``equal'', ``related'', and ``unrelated''.

\begin{equation}
\resizebox{0.9\hsize}{!}{$%
	MCC = \frac{(TP.FN) - (FP.FN)}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}\longrightarrow [-1,1]
	$%
}%
\end{equation}
MCC provides a fair measure of the ability with which a classifier can correctly predict both positive and negative instances \cite{chicco2020advantages}. MCC score of 0 represents random classification, however, with an increase in the number of true positives and true negatives, MCC moves closer to 1. It also takes into account the false positives and false negatives, which shift the mcc score towards -1. This measurement is markedly different from accuracy, that fails to account for imbalanced datasets and $F_1$ measure which is not affected by the true negative scores. As a result, MCC provides an acceptable alternate, in our current scenario comprising of imbalanced dataset (largely in favour of class ``unrelated'') to measuring the true performance of the models, used for threshold selection and model evaluation. Finally on a test dataset we evaluated our multi-class classification approach, using MCC and Cohen's Kappa coefficient ($\kappa$) \cite{gwet2014handbook} to identify the relationships between each pair of attributes.

\section{Experimental Setup}
\label{experimentalSetup}
In our earlier work \cite{Satti2020} $s_1$, $s_2$, and $s_5$ were used to generate over 115 million patient records, which are converted into a semi-structured form and stored in Hadoop Distributed File System (HDFS). We extended the same setup to create an additional 100,000 records, for 1000 patients with 3 medical fragments for $s_1$, $s_2$, and $s_4$, and 97 randomly selected and generated medical fragments amongst $s_1$, $s_2$, $s_3$, $s_4$ and $s_5$. These fragments, follow various design elements, supporting a variety of valid relational storage architectures. Such as, $s_1$, $s_2$ and $s_4$ are represented by creating a separate medical fragment for each participating table, $s_3$ utilizes its medical fragment to generate a linked record (from a linked object graph), where by the attributes can refer to other objects besides the elements of $t$, mimicking the application of explicit foreign keys, and $s_5$ is a flat table structure. The code to generate this data set is available at ``uhp\_map\_generation''\footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhpr\_storage}. This application produces three custom formatted files, containing an index for patients, an index for their medical fragments, and the medical fragment, corresponding to the EMR data. Using the medical fragments file, we then generate the semantically enriched attribute \footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhp\_map\_generation}, which contains the suffixes and their concepts corresponding to each EMR data attribute. The resulting set of enriched attributes are temporarily stored in a JSON file, which is then read by the same application to partially generate the schema maps. This process, is used to create 20,349 distinct pairs of attributes, across $s$. Each pair also contains the ``relationshipList'', which stores the results of fuzzy string matching\cite{FuzzyWuzzy} \footnote{Java Library: https://github.com/xdrop/fuzzywuzzy} between the attribute names. 
The JSON file thus produced, is then used by a python script to generate the semantically enriched sentences and their embedded vectors using Word2Vec, and 10 pre-trained BERT NLI models \cite{reimers-2019-sentence-bert}. The BERT models include ``bert-base-nli-stsb-mean-tokens'', ``bert-large-nli-stsb-mean-tokens'', ``roberta-base-nli-stsb-mean-tokens'', ``roberta-large-nli-stsb-mean-tokens'', ``distilbert-base-nli-stsb-mean-tokens'', ``bert-base-nli-mean-tokens'', ``bert-large-nli-mean-tokens'', ``roberta-base-nli-mean-tokens'', ``roberta-large-nli-mean-tokens'', and ``distilbert-base-nli-mean-tokens''. 
The embedding vectors are then compared using cosine similarity, which produces a score between -1 and 1.
The rationale behind switching the applications at various stages, is to cache the results and create checkpoints for restarting any failed stages, easily. Additionally, since python provides better support for easy generation of embedding vectors, it was thus preferred over the Java based implementation, which is otherwise very beneficial for other tools.
%These applications were executed on a workstation running Ubuntu 20.04.2 on top of AMD Ryzen 3 2200G, and 32GB ram.%

@@

The validity of our proposed approach has been evaluated using several techniques, including comparison of the proposed semantic matching process with fuzzy string matching, embedded vector generation and comparison using Word2Vec, and 10 BERT nli models.

@@

\subsection{Dataset Annotation}
In order to compare our computed models with ground truth, and to identify the best thresholds for classifying each instance as ``equal'', ``related'', or ``unrelated'' 4 human annotators were utilized, to anonymously, score the similarity of each pair of attribute names. In order to support this process, we first repurposed one of our generated data matrix, by marking all attribute pairs belonging to the same schema with the symbol ``-''. Following this, the annotators, marked each cell corresponding to a pair of attributes (conversely, each attribute pair corresponds to two cells, with the positioning of the pair-participants swapped; which is used for clarity and identify correct relationships between the attribute on left and attribute on right), by determining the similarity in terms of dissimilar as ``0'', exactly similar as ``1'', row attribute as child of column attribute as ``<'', row attribute as a parent of the column attribute as ``>'', and finally, unknown as ``~''. The data sheets generated after this extensive human effort, have been made available for other researchers\footnote{https://github.com/desertzebra/EMR-Interoperability/tree/master/Implemenation/Data/Annotated}.

@@

	\caption{One vs all binary classifications results}

@@

		\textbf{Model}                                        & \textbf{Class Positive} & \textbf{Class Negative} & \multicolumn{1}{c|}{\textbf{accuracy}} & \multicolumn{1}{c|}{\textbf{precision}} & \multicolumn{1}{c|}{\textbf{recall}} & \multicolumn{1}{c|}{\textbf{f-1}} & \multicolumn{1}{c|}{\textbf{mcc}} \\ \hline

>>>>

\begin{keywords}
Context aware systems, Decision support systems, Electronic medical records,Experts systems,   Health information management, Medical information systems, Ontology engineering, Text processing, Unsupervised learning
\end{keywords}

@@

Data interoperability resolves the heterogeneity between data artificats to enable seamless and interpretable communcation among source and target organizations, while preserving the data's original intention during storage, communication, and usage (as defined by IEEE 610.12 \cite{geraci1991ieee}, Health Level Seven International (HL7), and Healthcare Information and Management Systems Society HIMSS \cite{Himss_url2013}).
On the other hand, process interoperability regulates the communication among organizational processes to provide compatability between process artifacts within and seamless transformations across different organizations\cite{khan2013process}. Lastly, knowledge interoperability provides a sharing mechanism for reusing interpretable medical knowledge, acquired through expert intervention and other mechanisms, across decision support systems \cite{ali2017multi}.

@@

@@

Healthcare Standards such as HL7 - Fast Healthcare Interoperability Resources (FHIR), and openEHR provide the foundations for storing and communicating medical data, through the use of well defined protocols. While Systematized Nomenclature of Medicine—Clinical Terms (Snomed-CT) \cite{snomedct_url} and Logical Observation Identifiers Names and Codes (LOINC) \cite{loinc} provide a standard definition for clinical terminologies and laboratory tests, respectively. Similarly Medical Logic Module (MLM) provides a standardized way for expressing medical knowledge. However, the plethora of standards necessitates the creation of bridging standards that can resolve the heterogeniety between the medical standards. Substantial effort has gone into this endeavor with the Clinical Information Modeling Initiative (CIMI) \cite{CIMI2015} taking the lead in bridging the gap between HL7v3 and openEHR. Similarly, SNOMED CT and LOINC are working to resolve the redundancies between the two terminological standards since 2013. This healthcare interoperability solution follows a formal, albiet long process, which is greatly dependent on the human factor. However, the current healthcare scenario requires a quick solution to create a scaffolding of an interoperable bridge between various healthcare providers. It is also important to ensure that this scaffolding should be able to support the formal standardization processes of the future. In \cite{Satti2020}, we have presented the Ubiquitous Health Platform (UHP), which provides semantic reconciliation-on-read based data curation for resolving data interoperability between various schema. This methodology is based on the creation and management of schema maps, that can provide the framework for transforming a source schema into a target schema. 
In the current study, we will present our research work to build and manage the schema map knowledge base. Overall, our methodology is based on the creation, evaluation, and application of a novel schema matching technique to identify the relationships between attributes of the participating medical data schema. Since the terms used to identify attributes in the data schemas are not defined in any standard way, it is important to first identify the component words of the attribute term and then to append semantic concepts with these to create a meaningful sentence. This process is based on word expansion using concept lookup from Unified Medical Language System (UMLS). Once the sentence has been created, it is then trivial to create its embedded vectors representation using transformer based pre-trained models. The cosine similarity of any two embedded vectors can then indicate the degree of similarity between the original attributes. The motivation behind this methodology along with its development and evaluation details will be presented in the following flow:

@@

\begin{figure*}[t!]
	\centering
	\includegraphics[scale=0.60]{methodology}
	\caption{Methodology for creating schema maps.}
	\label{fig:methodology}
\end{figure*}

\section{Methodology}
\label{methodology}
Healthcare interoperability, with a focus on non-standard compliant medical schema is dependent on the generation and validation of schema maps as discussed above. To this end, the creation of a cohesive workflow is of utmost importance. In our earlier work \cite{Satti2020} we used maximum sequence identification and suffix tress based matching for syntactic matching of two distinct data schemas. This was followed by semantic concept enrichment and subsequently concept matching for creating rules in the form of schema maps. The simplified mapping functions, thereby created, provided a simple methodology for converting semi-structured medical data into an interpretable model form.
In our current methodology as visualized in Figure \ref{methodology}, we have utilized state-of-the-art natural language processing (NLP) techniques to extract the schema mapping rules from semi-structured data schemas. This methodology is based on identifying similarity between vector representations of two attributes, belonging to different medical schemas. 
Traditional NLP techniques such as Word2Vec are able to convert a word into an embedded vector, while Bidirectional Encoder Representations from Transformers (BERT) extracts an embedded vector from a sentence \cite{zhu2018exploring}. However, the terms forming the attribute names are bigger than a word (combination of multiple words) and smaller than a sentence. In order to resolve this problem, we extracted the set of suffixes from the terms forming the attribute names. The bidirectional nature of BERT, allows the creation of contextual embedded vectors, where each target word is affected by its neighboring words. Hence to convert the set of suffixes into a sentence, we collected the set of concepts corresponding to each suffix, from UMLS. This operation has two effects, firstly it is used to remove any suffix, which does not have a corresponding concept and secondly the extracted concepts are used to add context to each suffix and produce a contextual sentence. 
The following subsections provide the practical details for our methodology from schema acquisition to attribute name expansion, and finally schema map generation.

\subsection{Schema acquisition}
\label{schema_acq}
In the first step of our semantic reconciliation methodology, we simulate medical data acquisition from five distinct Electronic Medical Records (EMR) storage systems ($S$). These include patient reports from OpenEMR ($s_1$), 100,000 patient records from EMRBOTS ($s_2$) \cite{kartoun2016methodology}, custom database design by Pan et. al($s_3$) for supporting regional clinics and health care centers in China \cite{pan2016design}, clinical knowledge discovery tool MedTAKMI-CDI ($s_4$) \cite{inokuchi2007medtakmi}, and our custom implementation ($s_5$) \ref{ali2017reconciliation}. Each of these medical systems as shown in \ref{fig:schema}, follows the relational database design with logical entities such as demographics, diagnosis, medicine or others, placed into tables which can be further linked to one or more tables. While the database design implemented by each of these systems, fulfills the need of their respective information processing applications, the lack of interoperability, in terms of identifying similar attributes or exchanging the medical data is very much evident here.

@@

A similar notion of data heterogeneity, in terms of medical data schema is evident across the healthcare domain. This is caused by various factors, including the lack of one all-encompassing, and universally applicable terminological standard and different normalization level for representing attributes.\\
In the former case, while SNOMED-CT provides a mechanism for identifying the standard codes for clinical terms and LOINC can be used for laboratory related terms, most attribute names are created based on the gut feeling of the database designer. Additionally, while these codes can be used to represent data instances, the data schema, achieves no benefit from the same.  Consider the terms ``name'' and ``patientName'', which refer to the same attribute of the patient entity.  However, since there is no standard way to represent this attribute, both are considered correct ($s_1$ and $s_3$ use the former representation, while $s_2$ and $s_5$ use the latter).\\
In the later case, differences in normalization cause semantic differences, due to which some data could be available in one schema but absent in others such as OpenEMR demographics identifying the patient's residential location using specific attributes like ``Address'', ``City'', ``State'', ``Postal Code'', ``Country'', and others. Similarly, ``EncounterDate'' from $s_5$ is semantically similar to ``BeginDate'' of ``openemr\_MedicalProblems'' table in $s_1$, ``AdmissionStartDate'' of ``LabsCorePopulatedTable'' in $s_2$, 
``time'' in ``Diagnosis'' table of $s_3$, and  ``dateOfAdmission'' in ``Diagnosis'' and ``CareHistory'' tables of $s_4$.
Finally, $s_1$ and $s_3$ have separate tables containing the medicinal prescription details, however the same details are unavailable in $s_2$, $s_4$, and $s_5$. Once again, this is not an incorrect behavior since this information, might not be a part of the context or the requirements for the EMR/EHR storage systems.
In fact, the change in context of the medical data storage system from the initial time of development to a later stage of collaborative processing systems, is the main cause of heterogeneity. In order to provide an interoperable solution, it is therefore necessary to enhance the semantics of each data attribute by its contextually equivalent sentence.

\subsection{Attribute to Sentence Transformation}
\label{schema_prepro}
In order to process the EMR/EHR schema set $S$ and produce a set of corresponding semantically enriched sentences, we use the data representation $s_i$, generated through the process explained in sequence acquisition to collect the various medical fragments in memory. We then iterate over these fragments, building a set of attributes, distinguished by their name, schema's name, table's name, schema's version, source, and recorded data. This entails that ``PatientID'' from each of the four tables in $s_2$, and ``patientID'' from five tables in $s_4$, would result into nine attributes (assuming, as in the current case of no differences in versions of these systems).
For each attribute, we then generate the suffix array, which provides all possible substring representations contained within the attribute name. In order to generate the set of suffixes, we employ three strategies, forward suffix generation, whereby for a word $w$ of length $n$, $n-1$ suffixes of size 2 to $n-1$ are produced, backward suffix generation, to produce $n-1$ suffixes in reverse order with size $n-1$ to 2, and regular expression based suffix generation, which splits each word on, change of case, special characters(such as -, \_, !, and others), and numbers. In this way a large list of suffixes is generated, which is combined using a ``TreeSet'' data structure of Java, which internally sorts this list as well. An example of this suffix generation process, using the attribute name ``dateOfAdmission'' as it appears in $s_4$ is shown in Figure \ref{fig:suffix_attribute}.

\begin{figure}[tb!]
	\centering
	\includegraphics[scale=0.65]{suffixAttribute}
	\caption{An example of suffix arrays produced for the attribute ``dateOfAdmission''.}
	\label{fig:suffix_attribute}
\end{figure}

Suffix strings for similar attributes such as ``AdmissionStartDate''. ``diseaseNameOnAdmission'', and ``AdmissionEndDate'' appear in $s_2$, produce many, synctactically similar suffixes, to the presented example. This process, is only able to generate syntactic suffixes, producing many incoherent and unrelated suffixes. In order to counter this problem, and to limit the list of suffixes within the domain, we then query UMLS with exact search strategy, looking for the existence of any concepts against each suffix. In case, no semantic concept is found for a particular suffix, it is removed from the final Suffix Array. On the other hand, if atleast one semantic concept is found against the queried suffix, it is retained. Meanwhile the process continues for the next attribute, then the next table, and finally the next system, till no further processing is possible. 
The set of suffixes and their corresponding concepts are then used to build the sentence, where by each concept corresponding to a suffix is appended next to the suffix. An example of the resultant sentence for the attribute ``DateOfAdmission'' is shown in Table \ref{tab:sentence_example}.

\begin{table}[]
	\centering
	\caption{Sentence created from the attribute name ``DateOfAdmission''}
	\label{tab:sentence_example}
	\begin{tabular}{|p{\linewidth}|}
		\hline
		Date Value type - Date date allergenic extract Date in time Data types - Date Date Fruit;Of SPI1 wt Allele SPI1 gene TAF1 wt Allele BRIP1 gene Within Degrees fahrenheit Oral contraception BRIP1 wt Allele;Da Displacement of abomasum dalton Anterior descending branch of left coronary artery deca units cytarabine/daunorubicin protocol Dai Chinese Asymptomatic diagnosis of Drug Accountability Domain;ion Iontophoresis Route of Drug Administration Ions;on SPARC wt Allele Osteonectin SPARC gene On (qualifier value) Upon - dosing instruction fragment;Admission Admission activity Hospital admission;Dat SLC6A3 gene SLC6A3 wt Allele dopamine transporter Direct Coombs test SLC6A3 protein, human Test Date cytarabine/daunorubicin/thioguanine Alzheimer's Disease;mission Religious Missions;\\
		\hline
	\end{tabular}
\end{table}

@@

Schema Maps provide an interoperable bridge between two medical systems ($s_i \wedge s_j$), by identifying the semantic relationship between their participating attributes. This identification is based on the similarity between the embedded vectors, of the semantically enriched sentences corresponding to each data attribute. While the embedded vectors can be generated using any methodology, we tested 11 methodologies with Word2Vec and 10 models based on BERT (including BERT base, BERT large, distilBert, and RoBERTa) further detailed in section \ref{experimentalSetup}, the large/STSb version of RoBERTa\cite{reimers-2019-sentence-bert}, provides the best results. The pair of embedded vectors thus produced are then used to calculate cosine similarity, which is based on the inverse cosine distance between them. For our classification, we used the raw results (unnormalized) of cosine similarity, which produces a score between -1, and 1. Cosine similarity score of 0 indicates orthogonal relationship between the two vectors, which in our scenario indicates that the two sentences, and by extension their attributes are not related to each other. -1 indicates inverse relationship between the attributes, while 1 indicates the two attributes are very much the same.
For producing our schema maps, we are interested in three types of relationships, ``equal'' (the two attributes are same), ``related'' (the two attributes are related to each other), and ``unrelated'' (no relationship between the attributes). In order to classify the similarity results into one of these three classes, we then calculated the best thresholds using Matthews Correlation coefficient (MCC) \cite{chicco2020advantages} for classifying each instance as ``equal'', ``related'', and ``unrelated''.

\begin{equation}
\resizebox{0.9\hsize}{!}{$%
	MCC = \frac{(TP.FN) - (FP.FN)}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}\longrightarrow [-1,1]
	$%
}%
\end{equation}
MCC provides a fair measure of the ability with which a classifier can correctly predict both positive and negative instances \cite{chicco2020advantages}. MCC score of 0 represents random classification, however, with an increase in the number of true positives and true negatives MCC moves closer to 1. It also takes into account the false positives and false negatives, which shift the mcc score towards -1. This measurement is markedly different from accuracy that fails to account for imbalanced datasets and $F_1$ measure which is not affected by the true negative scores. As a result, MCC provides an acceptable alternate in our current scenario comprising of imbalanced dataset (largely in favour of class ``unrelated'') to measuring the true performance of the models, used for threshold selection and model evaluation. Finally on a test dataset we evaluated our multi-class classification approach using MCC and Cohen's Kappa coefficient ($\kappa$) \cite{gwet2014handbook} to identify the relationships between each pair of attributes.

\section{Experimental Setup}
\label{experimentalSetup}
In our earlier work \cite{Satti2020} $s_1$, $s_2$, and $s_5$ were used to generate over 115 million patient records, which are converted into a semi-structured form and stored in Hadoop Distributed File System (HDFS). We extended the same setup to create an additional 100,000 records, for 1000 patients with 3 medical fragments for $s_1$, $s_2$, and $s_4$, and 97 randomly selected and generated medical fragments amongst $s_1$, $s_2$, $s_3$, $s_4$ and $s_5$. These fragments follow various design elements supporting a variety of valid relational storage architectures. Such as, $s_1$, $s_2$ and $s_4$ are represented by creating a separate medical fragment for each participating table, $s_3$ utilizes its medical fragment to generate a linked record (from a linked object graph), where by the attributes can refer to other objects besides the elements of $t$, mimicking the application of explicit foreign keys, and $s_5$ is a flat table structure. The code to generate this data set is available at ``uhp\_map\_generation''\footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhpr\_storage}. This application produces three custom formatted files containing an index for patients, an index for their medical fragments, and the medical fragment corresponding to the EMR data. Using the medical fragments file, we then generate the semantically enriched attribute \footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhp\_map\_generation}, which contains the suffixes and their concepts corresponding to each EMR data attribute. The resulting set of enriched attributes are temporarily stored in a JSON file, which is then read by the same application to partially generate the schema maps. This process is used to create 20,349 distinct pairs of attributes across $s$. Each pair also contains the ``relationshipList'', which stores the results of fuzzy string matching\cite{FuzzyWuzzy} \footnote{Java Library: https://github.com/xdrop/fuzzywuzzy} between the attribute names. 
The JSON file thus produced is then used by a python script to generate the semantically enriched sentences and their embedded vectors using Word2Vec, and 10 pre-trained BERT NLI models \cite{reimers-2019-sentence-bert}. The BERT models include ``bert-base-nli-stsb-mean-tokens'', ``bert-large-nli-stsb-mean-tokens'', ``roberta-base-nli-stsb-mean-tokens'', ``roberta-large-nli-stsb-mean-tokens'', ``distilbert-base-nli-stsb-mean-tokens'', ``bert-base-nli-mean-tokens'', ``bert-large-nli-mean-tokens'', ``roberta-base-nli-mean-tokens'', ``roberta-large-nli-mean-tokens'', and ``distilbert-base-nli-mean-tokens''. 
The embedding vectors are then compared using cosine similarity, which produces a score between -1 and 1.
The rationale behind switching the applications at various stages is to cache the results and create checkpoints for restarting any failed stages, easily. Additionally, since python provides better support for easy generation of embedding vectors, it was thus preferred over the Java based implementation, which is otherwise very beneficial for other tools.
%These applications were executed on a workstation running Ubuntu 20.04.2 on top of AMD Ryzen 3 2200G, and 32GB ram.%

@@

The validity of our proposed approach has been evaluated using several techniques including comparison of the proposed semantic matching process with fuzzy string matching, embedded vector generation and comparison using Word2Vec, and 10 BERT nli models.

@@

\subsection{Dataset Annotation}
In order to compare our computed models with ground truth and to identify the best thresholds for classifying each instance as ``equal'', ``related'', or ``unrelated'' 4 human annotators were utilized to anonymously, score the similarity of each pair of attribute names. In order to support this process, we first repurposed one of our generated data matrix by marking all attribute pairs belonging to the same schema with the symbol ``-''. Following this, the annotators marked each cell corresponding to a pair of attributes (conversely, each attribute pair corresponds to two cells with the positioning of the pair-participants swapped; which is used for clarity and identify correct relationships between the attribute on left and attribute on right), by determining the similarity in terms of dissimilar as ``0'', exactly similar as ``1'', row attribute as child of column attribute as ``<'', row attribute as a parent of the column attribute as ``>'', and finally, unknown as ``~''. The data sheets generated after this extensive human effort have been made available for other researchers\footnote{https://github.com/desertzebra/EMR-Interoperability/tree/master/Implemenation/Data/Annotated}.

@@

	\caption{Performance matrix for individual classes using one vs all binarization technique}

@@

		\textbf{Model}                                        & \textbf{Class Positive} & \textbf{Class Negative} & \multicolumn{1}{c|}{\textbf{Accuracy}} & \multicolumn{1}{c|}{\textbf{Precision}} & \multicolumn{1}{c|}{\textbf{Recall}} & \multicolumn{1}{c|}{\textbf{F-1}} & \multicolumn{1}{c|}{\textbf{MCC}} \\ \hline
